---
title: "CS 422 Homework 6"
author: "Xingli Li"
output:
  html_notebook:
    toc: yes
    toc_float: yes
  html_document:
    toc: yes
    df_print: paged
---

## Homework 6

### Part 2

set work space and install dependent package.

```{r}

library(rpart)  #classification and regression trees
library(arules)
library(caret)
work.path <- getwd()
setwd(work.path)
```

### 2.1-(a)

Loading origin data and processing data.

...

### 2.1-(a)

Analysis data:

```{r}

transactions = c("1k","5k","20k","75k")

df <- data.frame(X = 0,
                 P1 = "",
                 P2 = "",
                 P3 = "",
                 P4 = "",
                 P5 = "",
                 P6 = "",
                 P7 = "",
                 P8 = "") # Prime the results data frame.

for (trx_id in transactions) {
   filename = paste0("tr-",trx_id,"-canonical.csv")
   df.temp <- read.csv(file = filename, encoding = "UTF-8")
   df <- rbind(df, df.temp)
   df.temp <- NULL
}

df <- df[-1,]

df
```

### 2.1-(b)

finding the frequent itemsets

discovering strong association rules within them.

### 2.1-(c)

Compare the rules you obtained for each different subset (1,000 -- 75,000 transactions). How does the number of transactions affect the results you observed? (Write the answer in your R markup file, easily identified.)

### 2.1-(d)-(i)

for the 75,000 transactions dataset using the same support level as determined in (b):

What is the most frequently purchased item or itemset?

filename \<- NULL

### 2.1-(d)-(ii)

What is the least frequently purchased item or itemset?
