---
title: "CS 422 Homework 6"
author: "Xingli Li"
output:
  html_notebook: 
    toc: yes
    toc_float: yes
    toc_depth: 5
  html_document:
    toc: yes
    df_print: paged
---

## Homework 6

### Part 2

set work space and install dependent package.

```{r}

library(rpart)  #classification and regression trees
library(arules)
library(arulesViz)
library(caret)
work.path <- getwd()
setwd(work.path)
```

### 2.1-(a)

Loading origin data and processing data.

### 2.1-(b)

Finding the frequent itemsets:

```{r}
# Read the data
trans <- read.transactions("tr-1k-canonical.csv",format = "basket", sep=",")
summary(trans)

### finding the frequent itemsets
frequent_itemsets <- apriori(trans, 
                             parameter=list(support=0.03, target="frequent itemsets"))
inspect(frequent_itemsets)

```

Discovering strong association rules within them.

```{r}
### get rules
parameter <- list(support=0.03,confidence = 0.9, target="rules")
rules <- apriori(trans, parameter=parameter)
#### Let's inspect the rules, sorted by confidence
rules.sorted_confd <- arules::sort(rules, decreasing = T,by = "confidence")

### Discovering the first 5 strong association rules within them.
inspect(head(rules.sorted_confd[1:5]))

```

```{r}
trans <- read.transactions("tr-5k-canonical.csv",sep=",")

### get rules
parameter <- list(support=0.03,confidence = 0.9, target="rules")
rules <- apriori(trans, parameter=parameter)
#### Let's inspect the rules, sorted by confidence
rules.sorted_confd <- arules::sort(rules, decreasing = T,by = "confidence")

### Discovering the first 5 strong association rules within them.
inspect(head(rules.sorted_confd[1:5]))

```

```{r}
trans <- read.transactions("tr-20k-canonical.csv",sep=",")

### get rules
parameter <- list(support=0.03,confidence = 0.9, target="rules")
rules <- apriori(trans, parameter=parameter)
#### Let's inspect the rules, sorted by confidence
rules.sorted_confd <- arules::sort(rules, decreasing = T,by = "confidence")

### Discovering the first 5 strong association rules within them.
inspect(head(rules.sorted_confd[1:5]))

```


```{r}
trans <- read.transactions("tr-20k-canonical.csv",sep=",")

### get rules
parameter <- list(support=0.03,confidence = 0.9, target="rules")
rules <- apriori(trans, parameter=parameter)
#### Let's inspect the rules, sorted by confidence
rules.sorted_confd <- arules::sort(rules, decreasing = T,by = "confidence")

### Discovering the first 5 strong association rules within them.
inspect(head(rules.sorted_confd[1:5]))

```


### 2.1-(c)

Compare the rules you obtained for each different subset (1,000 -- 75,000 transactions). How does the number of transactions affect the results you observed? (Write the answer in your R markup file, easily identified.)

```{r}
trans <- read.transactions("tr-1k-canonical.csv",sep=",")

### get rules
parameter <- list(support=0.03,confidence = 0.9, target="rules")
rules <- apriori(trans, parameter=parameter)
#### Let's inspect the rules, sorted by confidence
rules.sorted_confd <- arules::sort(rules, decreasing = T,by = "confidence")

### Discovering the first 5 strong association rules within them.
inspect(head(rules.sorted_confd[1:5]))
```

```{r}
trans <- read.transactions("tr-5k-canonical.csv",sep=",")

### get rules
parameter <- list(support=0.03,confidence = 0.9, target="rules")
rules <- apriori(trans, parameter=parameter)
#### Let's inspect the rules, sorted by confidence
rules.sorted_confd <- arules::sort(rules, decreasing = T,by = "confidence")

### Discovering the first 5 strong association rules within them.
inspect(head(rules.sorted_confd[1:5]))

```

```{r}
trans <- read.transactions("tr-20k-canonical.csv",sep=",")

### get rules
parameter <- list(support=0.03,confidence = 0.9, target="rules")
rules <- apriori(trans, parameter=parameter)
#### Let's inspect the rules, sorted by confidence
rules.sorted_confd <- arules::sort(rules, decreasing = T,by = "confidence")

### Discovering the first 5 strong association rules within them.
inspect(head(rules.sorted_confd[1:5]))
```

```{r}
trans <- read.transactions("tr-75k-canonical.csv",sep=",")

### get rules
parameter <- list(support=0.03,confidence = 0.9, target="rules")
rules <- apriori(trans, parameter=parameter)
#### Let's inspect the rules, sorted by confidence
rules.sorted_confd <- arules::sort(rules, decreasing = T,by = "confidence")

### Discovering the first 5 strong association rules within them.
inspect(head(rules.sorted_confd[1:5]))
```
Answer:
    With the increase of transaction data, the computation process takes more and more time.

### 2.1-(d)-(i)

For the 75,000 transactions dataset using the same support level as determined in (b):
What is the most frequently purchased item or itemset?

```{r}
trans <- read.transactions("tr-75k-canonical.csv",sep=",")

### finding the frequent itemsets
frequent_itemsets <- apriori(trans, 
                             parameter=list(support=0.03, target="frequent itemsets"))

inspect(frequent_itemsets)

summary(frequent_itemsets)

```

Answer:
    most frequent items: Hot Coffee, Coffee Eclair, Almond Twist, Apple Pie Apricot, Croissant

### 2.1-(d)-(ii)

What is the least frequently purchased item or itemset?

```{r}
data_df <- as(frequent_itemsets,"data.frame")

data_df[which.min(data_df$count),] 
```

Answer: 
    The least frequently purchased item is {Almond Twist,Hot Coffee}.
